HO unification from object language to meta language

### 2024-03-21 paper/main.tex
renaming
w
wip

\begin{proposition}[Simulation fidelity]\label{prop:fidelity}
In the context of~ \hrun, we have that
$\forall p \in 1 \ldots \mathcal{N}$
$$
\stepF{\mathcal{S}}{p}{\rho_{p-1}}{\rho_p}
\Leftrightarrow
\stepH{\mathcal{T}}{p}{\sigma_{p-1}}{\mathbb{L}}{\sigma_p}
$$
\end{proposition}
\noindent
In particular this property guarantees that a \emph{failure} in the \Fo{} run
is matched by a failure in \Ho{} \emph{at the same step}. We consider this
property very important from a practical point of view since it guarantees
that the execution traces are strongly related and in turn this enables a user
to debug a logic program in \Fo{} by looking at its execution trace in
\Ho{}.

@@

\begin{definition}[Subterm $\mathcal{P}(t)$]
$$
\begin{array}{l}
t \in \mathcal{P}(t)\\
t = f t_1\ldots t_n \Rightarrow \mathcal{P}(t_i) \subseteq \mathcal{P}(t) \land f \in \mathcal{P}(t)\\
t = \lambda x.t' \Rightarrow \mathcal{P}(t') \subseteq \mathcal{P}(t)\\
\end{array}
$$
\end{definition}

\noindent
We write $\mathcal{P}(X) = \bigcup_{t\in X} \mathcal{P}(t)$.

\newcommand{\wellb}{\ensuremath{\mathcal{W}}}
\begin{definition}[Well behaved set]
Given a set of terms $X \subseteq \Ho{}$,
$$
\wellb(X) \Leftrightarrow \forall t \in \mathcal{P}(X), t \not\in (\maybebeta{} ~\cup~ \maybeeta{})
$$
\end{definition}

@@

@@

\begin{elpicode}
type deref subst -> tm -> tm -> o.                   ~($\customlabel{pred:deref}{\sigma t}$)~
deref S (app A) (app B) :- map (deref S) A B.
deref S (lam F) (lam G) :- 
  pi x\ deref S x x => deref S (F x) (G x).
deref _ (con C) (con C).
deref S (uva N A) R :- set? N S F, move F A T, deref S T R.
deref S (uva N A) (uva N B) :- unset? N S, map (deref S) A B.
type move assignment -> list tm -> tm -> o.
move (abs Bo)        [H|L] R :- move (Bo H) L R.
move (val A)         []    A :- !.
move (val (uva N A)) L     (uva N X) :- std.append A L X.
\end{elpicode}

>>>>

\begin{proposition}[Simulation fidelity]\label{prop:fidelity}
In the context of~ \hrun, if $~\mathcal{T} \subseteq \llambda$ we have that
$\forall p \in 1 \ldots \mathcal{N}$
$$
\stepF{\mathcal{S}}{p}{\rho_{p-1}}{\rho_p}
\Leftrightarrow
\stepH{\mathcal{T}}{p}{\sigma_{p-1}}{\mathbb{L}}{\sigma_p}
$$
\end{proposition}
\noindent
In particular this property guarantees that a \emph{failure} in the \Fo{} run
is matched by a failure in \Ho{} \emph{at the same step}. We consider this
property very important from a practical point of view since it guarantees
that the execution traces are strongly related and in turn this enables a user
to debug a logic program in \Fo{} by looking at its execution trace in
\Ho{}.

XXX permuting hrun does not change the final result if check dooes not fail eagerly

XXX if we want to apply heuristics, we can apply them in decomp to avoid committing to
a non MGU too early

@@

\newcommand{\subterm}[1]{\ensuremath{\mathcal{P}(#1)}}
\begin{definition}[Subterm \subterm{t}]
$$
\begin{array}{l}
t \in \subterm{t}\\
t = f t_1\ldots t_n \Rightarrow \subterm{t_i} \subseteq \subterm{t} \land f \in \subterm{t}\\
t = \lambda x.t' \Rightarrow \subterm{t'} \subseteq \subterm{t}\\
\end{array}
$$
\end{definition}

\noindent
We write $\subterm{X} = \bigcup_{t\in X} \subterm{t}$.

\newcommand{\wellb}{\ensuremath{\mathcal{W}}}
\begin{definition}[Well behaved set]
Given a set of terms $X \subseteq \Ho{}$,
$$
\wellb(X) \Leftrightarrow \forall t \in \subterm{X}, t \not\in (\maybebeta{} ~\cup~ \maybeeta{})
$$
\end{definition}

@@

\subsection{Notations}

we use math mode for ho.

@@

\begin{elpicode}
type deref subst -> tm -> tm -> o.                   ~($\customlabel{pred:deref}{\sigma t}$)~
deref S (app A) (app B) :- map (deref S) A B.
deref S (lam F) (lam G) :- 
  pi x\ deref S x x => deref S (F x) (G x).
deref _ (con C) (con C).
deref S (uva N A) R :- set? N S F, move F A T, deref S T R.
deref S (uva N A) (uva N B) :- unset? N S, map (deref S) A B.
type move assignment -> list tm -> tm -> o.
move (abs Bo)        [H|L] R :- move (Bo H) L R.
move (val A)         []    A :- !.
move (val (uva N A)) L     (uva N X) :- append A L X.
\end{elpicode}
